Numerical Stability of Detonations in White Dwarf Simulations

### 2018-11-22 papers/ignition/paper.tex
Start describing the test problem

@@

%==========================================================================
% 1D collision test problem
%==========================================================================
\section{Numerical Test Problem}
\label{sec:collisions}

Plotfiles of relevant quantities are created every 0.01 seconds.
The timesteps are limited to ensure that every multiple of this plotfile interval is
reached exactly, which means an effective maximum timestep of 0.01 seconds. Timesteps
are by default limited only by the hydrodynamic stability constraint and this plotfile
interval constraint. For simulations with resolution better than 64 km, we use a CFL
number of 0.5. For the simulations with lower resolution, we use a CFL number of 0.01.
This is because at low resolution, the timestep is so large that the detonation can
form in a single timestep and represent such a large change that the burning integration
fails (\autoref{sec:retries}). The timestep retry mechanism can handle this gracefully,
but the subcycled timestep necessary to do so is usually so small that the full level
timestep ends up being so long that it just makes more sense to avoid this situation
entirely and use smaller level advance timesteps.

>>>>

%==========================================================================
% 1D collision test problem
%==========================================================================
\section{Test Problem}
\label{sec:collisions}

Our test problem is inspired by \cite{kushnir:2013}, and very loosely approximates the
conditions of two $0.64\ \msolar$ WDs colliding head-on. The simulation domain is 1D with a
reflecting boundary at $x = 0$. For $x > 0$ there is a uniform fluid composed (by mass)
of $50\%\, ^{12}$C, $45\%\, ^{16}$O, and $5\%\, ^{4}$He. The fluid is relatively cold,
$T = 10^7$ K, has density $\rho = 5 \times 10^6$ g/cm$^3$, and is traveling toward the
origin with velocity $-2 \times 10^8$ m/s. A uniform constant gravitational acceleration
is applied, $g = -1.1 \times 10^8$ m/s$^{2}$. This setup causes a sharp initial release
of energy at $x = 0$, and the primary question is whether a detonation occurs promptly
near this contact point, or occurs later (possibly at a distance from the contact point).
The simulated domain has width $1.6384 \times 10^9$ cm, and we apply inflow boundary conditions
that keep feeding the domain with material that has the same conditions as the initial fluid.
Simulations are performed with the adaptive mesh refinement (AMR) code \castro\
\citep{castro, astronum:2017}.

The only parameter we vary in this study is the spatial resolution used for this problem.
We vary both the base resolution of the grid, and also the amount of refinement allowed.
AMR is applied on gradients of the temperature: we tag zones for refinement if the temperature
varies by more than 50\% between two zones. Timesteps are limited only by the hydrodynamic
stability constraint, with CFL number 0.5. Although this leads to Strang splitting error
in the coupling of the burning and hydrodynamics for low resolution, we have verified that
the incorrect results seen at low resolution do not meaningfully depend on the timestep constraint
(both by applying a timestep limiter based on nuclear burning, and by using the spectral deferred
corrections driver in \castro, which directly couples the burning and hydrodynamics). At very high
resolution, the splitting error automatically becomes unimportant as the CFL criterion decreases
the timestep.

@@
